---
title: "Exponential Distribution compare to Central limit Theorem. Basic inferential data analysis"
author: "I D"
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
# set seed for reproducability, set lambda to 0.2, 40 samples, 1000 simulations
set.seed(31)
lambda <- 0.2
n <- 40
simulations <- 1000
```
# Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
# load the data ToothGrowth
data(ToothGrowth)
# preview the structure of the data
str(ToothGrowth)
# preview first 5 rows of the data
head(ToothGrowth, 5)
# Provide a basic summary of the data.
# data summary
summary(ToothGrowth)
# compare means of the different delivery methods
tapply(ToothGrowth$len,ToothGrowth$supp, mean)
```
```{r}
# plot data graphically
library(ggplot2)
ggplot(ToothGrowth, aes(factor(dose), len, fill = factor(dose))) +
  geom_boxplot() +
  # facet_grid(.~supp)+
  facet_grid(.~supp, labeller = as_labeller(
    c("OJ" = "Orange juice", 
      "VC" = "Ascorbic Acid"))) +
  labs(title = "Tooth growth of 60 guinea pigs by dosage and\nby delivery method of vitamin C",
       x = "Dose in milligrams/day", 
       y = "Tooth Lengh") +
  scale_fill_discrete(name = "Dosage of\nvitamin C\nin mg/day") +
  theme_classic()
```
```{r}
# Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
# comparison by delivery method for the same dosage
t05 <- t.test(len ~ supp, 
              data = rbind(ToothGrowth[(ToothGrowth$dose == 0.5) & 
                                         (ToothGrowth$supp == "OJ"),],
                           ToothGrowth[(ToothGrowth$dose == 0.5) & 
                                         (ToothGrowth$supp == "VC"),]), 
              var.equal = FALSE)
t1 <- t.test(len ~ supp, 
             data = rbind(ToothGrowth[(ToothGrowth$dose == 1) & 
                                        (ToothGrowth$supp == "OJ"),],
                          ToothGrowth[(ToothGrowth$dose == 1) & 
                                        (ToothGrowth$supp == "VC"),]), 
             var.equal = FALSE)
t2 <- t.test(len ~ supp, 
             data = rbind(ToothGrowth[(ToothGrowth$dose == 2) & 
                                        (ToothGrowth$supp == "OJ"),],
                          ToothGrowth[(ToothGrowth$dose == 2) & 
                                        (ToothGrowth$supp == "VC"),]), 
             var.equal = FALSE)
```
# summary of the conducted t.tests, which compare the delivery methods by dosage,
# take p-values and CI
```{r}
summaryBYsupp <- data.frame(
  "p-value" = c(t05$p.value, t1$p.value, t2$p.value),
  "Conf.Low" = c(t05$conf.int[1],t1$conf.int[1], t2$conf.int[1]),
  "Conf.High" = c(t05$conf.int[2],t1$conf.int[2], t2$conf.int[2]),
  row.names = c("Dosage .05","Dosage 1","Dosage 2"))
```
```{r}
# show data table 
summaryBYsupp
```
# Conclusion
# For dosage of .5 milligrams/day and 1 milligrams/day does matter the delivery method. the delivery method for 2 milligrams/day. For dosage of 2 milligrams/day the delivery method doesn’t matter.